---
title: "School Climate and Student Growth"
author: "Michael O'key, Michael Chrzan, Laisha Ozuna"
date: "`r Sys.Date()`"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Honor Code Statement

We strongly encourage students to form study groups and students may discuss and work on assignments in groups. We expect that each student understands their own submission. As such, students must write their submissions independently and clearly disclose the names of all other students who were part of their study group. Additionally, lifting code or solutions directly from the internet (e.g., Google, GitHub, Stack Overflow) is a violation of the [Stanford Honor Code](https://communitystandards.stanford.edu/policies-and-guidance/honor-code). We take academic honesty and Honor Code violations extremely seriously and expect the same of students. If you have questions about what may or may not constitute an Honor Code violation, please reach out the teaching team.

**Email: michael.okey\@stanford.edu, mlchrzan\@stanford.edu, lozuna\@stanford.edu**

**SUID: ???, 05745732, ???**

**Study Group: The authors**

I acknowledge and agree to abide by the Honor Code.

**Signed: Michael O'key, Michael Leon Chrzan, Laisha Ozuna**

```{r libraries, include=FALSE}
library(tidyverse)
library(tidytext)
library(stringr)
library(readxl)
library(curl)
library(janitor)
library(RColorBrewer)
library(viridis)
library(skimr)
library(cowplot)
library(ggrepel)
library(patchwork)
library(cartography)
library(here)
```

# Research Question

What relationship, if any, exists between gentrification, as operationalized by the Pearman and Greene (2022) model, and educational outcomes in San Francisco and Detroit between 2010 and 2020?

# Import Data

```{r housing-data, echo=FALSE, include=FALSE}
#ACS Estimate Data 
det2020 <- read_csv("Detroit Housing Data/Detroit housing Data 2020.csv") |> clean_names()
det2019 <- read_csv("Detroit Housing Data/Detroit housing Data 2019.csv") |> clean_names()
det2018 <- read_csv("Detroit Housing Data/Detroit housing Data 2018.csv") |> clean_names()
det2017 <- read_csv("Detroit Housing Data/Detroit housing Data 2017.csv") |> clean_names()
det2016 <- read_csv("Detroit Housing Data/Detroit housing Data 2016.csv") |> clean_names()
det2015 <- read_csv("Detroit Housing Data/Detroit housing Data 2015.csv") |> clean_names()
det2014 <- read_csv("Detroit Housing Data/Detroit housing Data 2014.csv") |> clean_names()
det2013 <- read_csv("Detroit Housing Data/Detroit housing Data 2013.csv") |> clean_names()
det2012 <- read_csv("Detroit Housing Data/Detroit housing Data 2012.csv") |> clean_names()
det2011 <- read_csv("Detroit Housing Data/Detroit housing Data 2011.csv") |> clean_names()
det2010 <- read_csv("Detroit Housing Data/Detroit housing Data 2010.csv") |> clean_names()

#Detroit Parcel Data
det_parcel <- read_csv("Detroit Housing Data/Parcels_filtered.csv") |> clean_names()

```

```{r acs-ed-data, echo=FALSE, include=FALSE}

acs_detroit_2010 <- read_csv('ACS Detroit Education Data/ACSDtrt2010-Data.csv') 
acs_detroit_2011 <- read_csv('ACS Detroit Education Data/ACSDtrt2011-Data.csv') 
acs_detroit_2012 <- read_csv('ACS Detroit Education Data/ACSDtrt2012-Data.csv') 
acs_detroit_2013 <- read_csv('ACS Detroit Education Data/ACSDtrt2013-Data.csv') 
acs_detroit_2014 <- read_csv('ACS Detroit Education Data/ACSDtrt2014-Data.csv') 
acs_detroit_2015 <- read_csv('ACS Detroit Education Data/ACSDtrt2015-Data.csv') 
acs_detroit_2016 <- read_csv('ACS Detroit Education Data/ACSDtrt2016-Data.csv') 
acs_detroit_2017 <- read_csv('ACS Detroit Education Data/ACSDtrt2017-Data.csv') 
acs_detroit_2018 <- read_csv('ACS Detroit Education Data/ACSDtrt2018-Data.csv') 
acs_detroit_2019 <- read_csv('ACS Detroit Education Data/ACSDtrt2019-Data.csv')
acs_detroit_2020 <- read_csv('ACS Detroit Education Data/ACSDtrt2020-Data.csv')

```

```{r nces-data, echo=FALSE, include=FALSE}
#Files named 'Combined' (from NCES download)
nces2010 <- read_csv('NCES Data/Combined/(2009-2010) Combined.csv') |> clean_names()
nces2011 <- read_csv('NCES Data/Combined/(2010-2011) Combined.csv') |> clean_names()
nces2012 <- read_csv('NCES Data/Combined/(2011-2012) Combined.csv') |> clean_names()
nces2013 <- read_csv('NCES Data/Combined/(2012-2013) Combined.csv') |> clean_names()
nces2014 <- read_csv('NCES Data/Combined/(2013-2014) Combined.csv') |> clean_names()
nces2015 <- read_csv('NCES Data/Combined/(2014-2015) Combined.csv') |> clean_names()
nces2016 <- read_csv('NCES Data/Combined/(2015-2016) Combined.csv') |> clean_names()
nces2017 <- read_csv('NCES Data/Combined/(2016-2017) Combined.csv') |> clean_names()
nces2018 <- read_csv('NCES Data/Combined/(2017-2018) Combined.csv') |> clean_names()

#Readiness Files
nces2016_readiness <- read_csv('NCES Data/Readiness/(2015-2016) College Readiness.csv') |> clean_names()
nces2017_readiness <- read_csv('NCES Data/Readiness/(2016-2017) College Readiness.csv') |> clean_names()
nces2018_readiness <- read_csv('NCES Data/Readiness/(2017-2018) College Readiness.csv') |> clean_names()

```

# Gentrification Data

## ACS Housing Data Cleaning

```{r clean-housing}
# List of years
years <- 2010:2020

# Loop over each year
for(year in years) {
  # Construct the tibble name
  tibble_name <- paste0("det", year)
  
  # Evaluate the tibble name to get the actual tibble
  tibble <- eval(parse(text = tibble_name))
  
  # Perform the operations
  tibble <- tibble |> 
    select(-starts_with("units_in_structure"),
           -starts_with("rooms"),
           -starts_with("bedrooms"),
           -starts_with("housing_tenure"),
           -starts_with("vehicles_available"),
           -starts_with("house_heating"),
           -starts_with("selected_characteristics"),
           -starts_with("occupants_per_room"), 
           -starts_with("year_householder"),
           -starts_with("selected_monthly_owner_costs"),
           -starts_with("mortgage_status"),
           -starts_with("year_structure_built"),
           -contains("grapi"),
           -housing_occupancy, 
           -value, 
           -mortgage_status,
           -selected_monthly_owner_costs_smoc,
           -gross_rent, 
           -selected_monthly_owner_costs_as_a_percentage_of_household_income_smocapi, 
           -selected_monthly_owner_costs_smoc,
           -gross_rent_as_a_percentage_of_household_income_grapi) |> 
    mutate(year = year)
  
  tibble <- tibble |> 
    mutate(label_grouping = str_trim(label_grouping)) |> 
    filter(label_grouping == 'Estimate') |> 
    select(-label_grouping)
  
  #Assign the modified tibble back to the original tibble
  assign(tibble_name, tibble, envir = .GlobalEnv)
}

det_data <- bind_rows(det2010, det2011, det2012,det2013, det2014, det2015, det2016, det2017, det2018, det2019, det2020) |> 
  select(year, everything())

det_data <- det_data |>
  unite('occupied_housing_units', contains('occupied_housing_units'), na.rm = TRUE) |> 
  unite('vacant_housing_units', contains('vacant_housing_units'), na.rm = TRUE) |> 
  unite('homeowner_vacancy_rate', contains('homeowner_vacancy_rate'), na.rm = TRUE)  |> 
  unite('rental_vacancy_rate', contains('rental_vacancy_rate'), na.rm = TRUE)  |> 
  unite("value_less_than_50_000", contains("less_than_50_000"), na.rm = TRUE) |> 
  unite("value_50_000_to_99_999", contains("50_000_to_99_999"), na.rm = TRUE) |> 
  unite('value_100_000_to_149_999', contains('100_000_to_149_999'), na.rm = TRUE) |> 
  unite('value_150_000_to_199_999', contains('150_000_to_199_999'), na.rm = TRUE) |> 
  unite('value_200_000_to_299_999', contains('200_000_to_299_999'), na.rm = TRUE) |> 
  unite('value_300_000_to_499_999', contains('300_000_to_499_999'), na.rm = TRUE) |> 
  unite('value_500_000_to_999_999', contains('500_000_to_999_999'), na.rm = TRUE) |> 
  unite('value_1_000_000_or_more', contains('1_000_000_or_more'), na.rm = TRUE) |> 
  unite('value_median_dollars', value_median_dollars, value_owner_occupied_units_median_dollars, na.rm = TRUE) |> 
  unite('gross_rent_median_dollars', gross_rent_median_dollars, gross_rent_occupied_units_paying_rent_median_dollars, na.rm = TRUE) |> 
  unite('gross_rent_no_rent_paid', contains('no_rent_paid'), na.rm = TRUE) |> 
  unite('gross_rent_1_500_or_more', gross_rent_1_500_or_more, gross_rent_occupied_units_paying_rent_1_500_or_more, na.rm = TRUE) |> 
  unite("gross_rent_less_than_200", contains('rent_less_than_200'), na.rm = TRUE) |>
  unite("gross_rent_200_to_299", contains('rent_200_to_299'), na.rm = TRUE) |> 
  unite("gross_rent_300_to_499", contains('rent_300_to_499'), na.rm = TRUE) |> 
  unite("gross_rent_500_to_749", contains('rent_500_to_749'), na.rm = TRUE) |> 
  unite("gross_rent_750_to_999", contains('rent_750_to_999'), na.rm = TRUE) |> 
  unite("gross_rent_1_000_to_1_499", contains('1_000_to_1_499'), na.rm = TRUE)
  

det_data <- det_data |>  
  mutate(across(where(is.character), ~str_replace_all(., ',', ''))) |> 
  mutate_all(as.numeric) |> 
  mutate(structures_built = case_when(
    year == 2014 ~ 6, 
    year == 2015 ~ 45, 
    year == 2016 ~ 173, 
    year == 2017 ~ 202, 
    year == 2018 ~ 467, 
    year == 2019 ~ 1115, 
    year == 2020 ~ 99))


det_data <- det_data |> 
  mutate(gross_rent_less_than_500 = gross_rent_less_than_200+gross_rent_200_to_299+gross_rent_300_to_499, 
         gross_rent_500_to_999 = gross_rent_500_to_749 + gross_rent_750_to_999, 
         gross_rent_1_500_or_more_occupied = gross_rent_occupied_units_paying_rent_1_500_to_1_999+gross_rent_occupied_units_paying_rent_2_000_to_2_499+gross_rent_occupied_units_paying_rent_2_500_to_2_999+gross_rent_occupied_units_paying_rent_3_000_or_more, 
         .keep = "unused")

det_data <- det_data |> 
  unite("gross_rent_less_than_500", contains('rent_less_than_500'), na.rm = TRUE) |> 
  unite("gross_rent_500_to_999", contains('rent_500_to_999'), na.rm = TRUE) |> 
  unite("gross_rent_1_500_or_more", contains('1_500_or_more'), na.rm = TRUE) 

det_data <- det_data |> 
  select(year, structures_built, contains('median'), starts_with('value'), gross_rent_less_than_500, gross_rent_500_to_999, gross_rent_1_000_to_1_499, gross_rent_1_500_or_more, starts_with('gross'), everything()) |> 
  mutate(across(everything(), as.numeric))
```

## ACS Education Data Cleaning

```{r clean-columns}
colnames(acs_detroit_2010) <- acs_detroit_2010[1, ]
acs_detroit_2010 <- acs_detroit_2010[-1, ]

colnames(acs_detroit_2011) <- acs_detroit_2011[1, ]
acs_detroit_2011 <- acs_detroit_2011[-1, ]

colnames(acs_detroit_2012) <- acs_detroit_2012[1, ]
acs_detroit_2012 <- acs_detroit_2012[-1, ]

colnames(acs_detroit_2013) <- acs_detroit_2013[1, ]
acs_detroit_2013 <- acs_detroit_2013[-1, ]

colnames(acs_detroit_2014) <- acs_detroit_2014[1, ]
acs_detroit_2014 <- acs_detroit_2014[-1, ]

colnames(acs_detroit_2015) <- acs_detroit_2015[1, ]
acs_detroit_2015 <- acs_detroit_2015[-1, ]

colnames(acs_detroit_2016) <- acs_detroit_2016[1, ]
acs_detroit_2016 <- acs_detroit_2016[-1, ]

colnames(acs_detroit_2017) <- acs_detroit_2017[1, ]
acs_detroit_2017 <- acs_detroit_2017[-1, ]

colnames(acs_detroit_2018) <- acs_detroit_2018[1, ]
acs_detroit_2018 <- acs_detroit_2018[-1, ]

colnames(acs_detroit_2019) <- acs_detroit_2019[1, ]
acs_detroit_2019 <- acs_detroit_2019[-1, ]

colnames(acs_detroit_2020) <- acs_detroit_2020[1, ]
acs_detroit_2020 <- acs_detroit_2020[-1, ]
```

```{r clean-variables}
#parse for educational attainment data
educ_detroit_2010 <- acs_detroit_2010 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
  clean_names()

educ_detroit_2011 <- acs_detroit_2011 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
  clean_names()

educ_detroit_2012 <- acs_detroit_2012 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
  clean_names()

educ_detroit_2013 <- acs_detroit_2013 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
  clean_names()

educ_detroit_2014 <- acs_detroit_2014 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
   clean_names()

educ_detroit_2015 <- acs_detroit_2015 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
   clean_names()

educ_detroit_2016 <- acs_detroit_2016 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
   clean_names()

educ_detroit_2017 <- acs_detroit_2017 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
   clean_names()

educ_detroit_2018 <- acs_detroit_2018 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
  clean_names()

educ_detroit_2019 <- acs_detroit_2019 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
   clean_names()

educ_detroit_2020 <- acs_detroit_2020 |>
  select(1:2, starts_with("Estimate!!EDUCATIONAL ATTAINMENT!!"), 
         starts_with("Percent!!EDUCATIONAL ATTAINMENT!!")) |>
   clean_names()

```

```{r convert-to-numeric}
convert_to_dbl <- function(x) {
  ifelse(str_detect(x, "^\\d+\\.?\\d*$"), as.numeric(x), x)
}

# apply <dbl> to select columns and add year
educ_detroit_2010 <- educ_detroit_2010 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2010) |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

educ_detroit_2011 <- educ_detroit_2011 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2011) |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

educ_detroit_2012 <- educ_detroit_2012 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2012) |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

educ_detroit_2013 <- educ_detroit_2013 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2013) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_less_than_9th_grade = "percent_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         percent_educational_attainment_9th_to_12th_grade_no_diploma = "percent_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         percent_educational_attainment_high_school_graduate_includes_equivalency = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         percent_educational_attainment_some_college_no_degree = "percent_educational_attainment_population_25_years_and_over_some_college_no_degree",
         percent_educational_attainment_associates_degree = "percent_educational_attainment_population_25_years_and_over_associates_degree",
         percent_educational_attainment_bachelors_degree = "percent_educational_attainment_population_25_years_and_over_bachelors_degree",
         percent_educational_attainment_graduate_or_professional_degree = "percent_educational_attainment_population_25_years_and_over_graduate_or_professional_degree") |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

educ_detroit_2014 <- educ_detroit_2014 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2014) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_less_than_9th_grade = "percent_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         percent_educational_attainment_9th_to_12th_grade_no_diploma = "percent_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         percent_educational_attainment_high_school_graduate_includes_equivalency = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         percent_educational_attainment_some_college_no_degree = "percent_educational_attainment_population_25_years_and_over_some_college_no_degree",
         percent_educational_attainment_associates_degree = "percent_educational_attainment_population_25_years_and_over_associates_degree",
         percent_educational_attainment_bachelors_degree = "percent_educational_attainment_population_25_years_and_over_bachelors_degree",
         percent_educational_attainment_graduate_or_professional_degree = "percent_educational_attainment_population_25_years_and_over_graduate_or_professional_degree") |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

educ_detroit_2015 <- educ_detroit_2015 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2015) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_less_than_9th_grade = "percent_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         percent_educational_attainment_9th_to_12th_grade_no_diploma = "percent_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         percent_educational_attainment_high_school_graduate_includes_equivalency = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         percent_educational_attainment_some_college_no_degree = "percent_educational_attainment_population_25_years_and_over_some_college_no_degree",
         percent_educational_attainment_associates_degree = "percent_educational_attainment_population_25_years_and_over_associates_degree",
         percent_educational_attainment_bachelors_degree = "percent_educational_attainment_population_25_years_and_over_bachelors_degree",
         percent_educational_attainment_graduate_or_professional_degree = "percent_educational_attainment_population_25_years_and_over_graduate_or_professional_degree") |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

educ_detroit_2016 <- educ_detroit_2016 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2016) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_less_than_9th_grade = "percent_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         percent_educational_attainment_9th_to_12th_grade_no_diploma = "percent_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         percent_educational_attainment_high_school_graduate_includes_equivalency = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         percent_educational_attainment_some_college_no_degree = "percent_educational_attainment_population_25_years_and_over_some_college_no_degree",
         percent_educational_attainment_associates_degree = "percent_educational_attainment_population_25_years_and_over_associates_degree",
         percent_educational_attainment_bachelors_degree = "percent_educational_attainment_population_25_years_and_over_bachelors_degree",
         percent_educational_attainment_graduate_or_professional_degree = "percent_educational_attainment_population_25_years_and_over_graduate_or_professional_degree") |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

#Converstion Continued (blocks combined into one)
educ_detroit_2017 <- educ_detroit_2017 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2017) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_less_than_9th_grade = "percent_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         percent_educational_attainment_9th_to_12th_grade_no_diploma = "percent_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         percent_educational_attainment_high_school_graduate_includes_equivalency = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         percent_educational_attainment_some_college_no_degree = "percent_educational_attainment_population_25_years_and_over_some_college_no_degree",
         percent_educational_attainment_associates_degree = "percent_educational_attainment_population_25_years_and_over_associates_degree",
         percent_educational_attainment_bachelors_degree = "percent_educational_attainment_population_25_years_and_over_bachelors_degree",
         percent_educational_attainment_graduate_or_professional_degree = "percent_educational_attainment_population_25_years_and_over_graduate_or_professional_degree") |>
  select(-estimate_educational_attainment_percent_high_school_graduate_or_higher, -estimate_educational_attainment_percent_bachelors_degree_or_higher)

educ_detroit_2018 <- educ_detroit_2018 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2018) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree") |>                            #add percentages
  mutate(percent_educational_attainment_population_25_years_and_over = estimate_educational_attainment_population_25_years_and_over,
         percent_educational_attainment_less_than_9th_grade = round((estimate_educational_attainment_less_than_9th_grade / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_9th_to_12th_grade_no_diploma = round((estimate_educational_attainment_9th_to_12th_grade_no_diploma / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_high_school_graduate_includes_equivalency = round((estimate_educational_attainment_high_school_graduate_includes_equivalency / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_some_college_no_degree = round((estimate_educational_attainment_some_college_no_degree / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_associates_degree = round((estimate_educational_attainment_associates_degree / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_bachelors_degree = round((estimate_educational_attainment_bachelors_degree / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_graduate_or_professional_degree = round((estimate_educational_attainment_graduate_or_professional_degree / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_percent_high_school_graduate_or_higher = round((estimate_educational_attainment_population_25_years_and_over_high_school_graduate_or_higher / estimate_educational_attainment_population_25_years_and_over) * 100, 1),
         percent_educational_attainment_percent_bachelors_degree_or_higher = round((estimate_educational_attainment_population_25_years_and_over_bachelors_degree_or_higher / estimate_educational_attainment_population_25_years_and_over) * 100, 1)) |>
  select(-estimate_educational_attainment_population_25_years_and_over_high_school_graduate_or_higher, -estimate_educational_attainment_population_25_years_and_over_bachelors_degree_or_higher)
  
educ_detroit_2019 <- educ_detroit_2019 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2019) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_less_than_9th_grade = "percent_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         percent_educational_attainment_9th_to_12th_grade_no_diploma = "percent_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         percent_educational_attainment_high_school_graduate_includes_equivalency = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         percent_educational_attainment_some_college_no_degree = "percent_educational_attainment_population_25_years_and_over_some_college_no_degree",
         percent_educational_attainment_associates_degree = "percent_educational_attainment_population_25_years_and_over_associates_degree",
         percent_educational_attainment_bachelors_degree = "percent_educational_attainment_population_25_years_and_over_bachelors_degree",
         percent_educational_attainment_graduate_or_professional_degree = "percent_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_percent_high_school_graduate_or_higher = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_or_higher",
         percent_educational_attainment_percent_bachelors_degree_or_higher = "percent_educational_attainment_population_25_years_and_over_bachelors_degree_or_higher"  
         ) |>
  select(-estimate_educational_attainment_population_25_years_and_over_high_school_graduate_or_higher, -estimate_educational_attainment_population_25_years_and_over_bachelors_degree_or_higher)

educ_detroit_2020 <- educ_detroit_2020 |>
  mutate_all(convert_to_dbl) |>
  mutate(year = 2020) |>
  rename(estimate_educational_attainment_less_than_9th_grade =
           "estimate_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         estimate_educational_attainment_9th_to_12th_grade_no_diploma = "estimate_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         estimate_educational_attainment_high_school_graduate_includes_equivalency = "estimate_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         estimate_educational_attainment_some_college_no_degree = "estimate_educational_attainment_population_25_years_and_over_some_college_no_degree",
         estimate_educational_attainment_associates_degree = "estimate_educational_attainment_population_25_years_and_over_associates_degree",
         estimate_educational_attainment_bachelors_degree = "estimate_educational_attainment_population_25_years_and_over_bachelors_degree",
         estimate_educational_attainment_graduate_or_professional_degree = "estimate_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_less_than_9th_grade = "percent_educational_attainment_population_25_years_and_over_less_than_9th_grade",
         percent_educational_attainment_9th_to_12th_grade_no_diploma = "percent_educational_attainment_population_25_years_and_over_9th_to_12th_grade_no_diploma",
         percent_educational_attainment_high_school_graduate_includes_equivalency = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_includes_equivalency",
         percent_educational_attainment_some_college_no_degree = "percent_educational_attainment_population_25_years_and_over_some_college_no_degree",
         percent_educational_attainment_associates_degree = "percent_educational_attainment_population_25_years_and_over_associates_degree",
         percent_educational_attainment_bachelors_degree = "percent_educational_attainment_population_25_years_and_over_bachelors_degree",
         percent_educational_attainment_graduate_or_professional_degree = "percent_educational_attainment_population_25_years_and_over_graduate_or_professional_degree",
         percent_educational_attainment_percent_high_school_graduate_or_higher = "percent_educational_attainment_population_25_years_and_over_high_school_graduate_or_higher",
         percent_educational_attainment_percent_bachelors_degree_or_higher = "percent_educational_attainment_population_25_years_and_over_bachelors_degree_or_higher"  
         ) |>
  select(-estimate_educational_attainment_population_25_years_and_over_high_school_graduate_or_higher, -estimate_educational_attainment_population_25_years_and_over_bachelors_degree_or_higher)
```

```{r bind-ed-acs-data}
educ_detroit <- bind_rows(educ_detroit_2010,
                          educ_detroit_2011,
                          educ_detroit_2012,
                          educ_detroit_2013,
                          educ_detroit_2014,
                          educ_detroit_2015,
                          educ_detroit_2016,
                          educ_detroit_2017,
                          educ_detroit_2018,
                          educ_detroit_2019,
                          educ_detroit_2020)
```

## Combining All ACS Data

```{r join-acs}
acs_all_data <- full_join(educ_detroit, det_data, 
                          by = 'year')

acs_all_data <- acs_all_data |> 
  select(year, everything())
```

# Education Outcomes Data

## NCES Data Cleaning

```{r join-combined-data}
#Converstion of all variables to a character value for joining
all_objects <- ls()

#Filter the list of objects based on the naming convention "ncesYYYY"
nces_objects <- all_objects[str_detect(all_objects, "^nces\\d{4}$")]

#Loop over the filtered list of objects
for (obj_name in nces_objects) {
  #Get the actual object from its name
  obj <- get(obj_name)
  
  #Check if the object is a tibble
  if (is_tibble(obj)) {
    
    #Convert all variables to character
    obj <- obj |> 
      mutate(across(everything(), as.character))
    
    #Filtering for DPSCD District Level Data 
    obj <- obj |>  
      filter(district_code == 82015 & entity_type == "LEA District") |>  
      mutate(year = substr(obj_name, 5, 8))
    
    #Assign the modified object back to its original name in the global environment
    assign(obj_name, obj, envir = .GlobalEnv)
  }
}

nces_all_combined <- bind_rows(nces2010,
                               nces2011, 
                               nces2012, 
                               nces2013, 
                               nces2014, 
                               nces2015, 
                               nces2016, 
                               nces2017, 
                               nces2018) 

nces_all_combined <- nces_all_combined |> 
  select(year, 
         contains('enrollment'), 
         contains('percent'), 
         contains('rate'), 
         -grade_11_reading_percent_ready, 
         -grade_11_science_percent_ready)

nces_all_combined
```

```{r join-readiness-data}
#Define the list of dataframes and years
nces_ready_dfs <- list(nces2016_readiness, nces2017_readiness, nces2018_readiness)
years <- c(2016, 2017, 2018)

#Initialize an empty list to store the results
results <- list()

#Loop over each dataframe
for(i in seq_along(nces_ready_dfs)) {
  # Perform the operations
  results[[i]] <- nces_ready_dfs[[i]] |> 
    filter(district_code == '82015' & building_name == 'All Buildings' & subgroup == 'All Students') |> 
    mutate(year = years[i]) |> 
    select(year, 
           math_percent_ready, 
           ebrw_percent_ready, 
           all_subject_percent_ready)
}

#Assign the results back to the original variables
nces2016_readiness <- results[[1]]
nces2017_readiness <- results[[2]]
nces2018_readiness <- results[[3]]

#Bind Years into One Tibble
nces_16_18_readiness <- bind_rows(nces2016_readiness, 
                                  nces2017_readiness, 
                                  nces2018_readiness)

nces_16_18_readiness
```

# Combining all NCES Data

```{r join-all-nces-data}

nces_16_18_readiness <- nces_16_18_readiness |> 
  mutate(year = as.character(year))

nces_all_data <- full_join(nces_all_combined, nces_16_18_readiness, 
          by = 'year') |> 
  unite("grade_11_math_percent_ready", 
        contains('math_percent_ready'), 
        na.rm = TRUE) |>
  unite("grade_11_english_percent_ready", 
        grade_11_english_percent_ready, ebrw_percent_ready,
        na.rm = TRUE) |>
  unite("grade_11_all_subject_percent_ready", 
        contains('all_subject_percent_ready'), 
        na.rm = TRUE) |> 
  mutate(grade_11_math_percent_ready = str_replace_all(grade_11_math_percent_ready, "0_", ""), 
         grade_11_english_percent_ready = str_replace_all(grade_11_english_percent_ready, "0_", ""), 
         grade_11_all_subject_percent_ready = str_replace_all(grade_11_all_subject_percent_ready, "0_", ""))

nces_all_data <- nces_all_data |> 
  mutate(across(everything(), as.numeric))

nces_all_data
```

# Construction Data

```{r parcel-data}

unique(det_parcel$property_class_desc)

housing_built <- det_parcel |> 
  filter(year_built >= 2009 & year_built < 2019) |> 
  group_by(year_built) |> 
  summarise(housing_built = n()) |> 
  ungroup() |> 
  rename(year = year_built) 

housing_built
```

# Combining All Data

```{r join-all-data}

acs_nces_data <- full_join(acs_all_data, nces_all_data, 
                           by = 'year') |> 
  filter(year < 2019) |> 
  select(-geographic_area_name, -geography)

acs_nces_data

all_data <- filter(housing_built, year > 2009) |> 
  full_join(acs_nces_data, housing_built, 
                      by = 'year')

all_data
```

# Calculating Gentrification

```{r calc-gentrication}
all_data_gentrify <- all_data |> 
  mutate(rent_delta = gross_rent_median_dollars - lag(gross_rent_median_dollars), 
         value_delta = value_median_dollars - lag(value_median_dollars), 
         edu_level_delta = percent_educational_attainment_percent_bachelors_degree_or_higher - lag(percent_educational_attainment_percent_bachelors_degree_or_higher), 
         housing_delta = housing_built - lag(housing_built)) |> 

  mutate(gentrifying = if_else((rent_delta > 0 | value_delta > 0) & edu_level_delta > 0 & housing_delta > 0, TRUE, FALSE)) |> 
  select(year, rent_delta, value_delta, edu_level_delta, housing_delta, gentrifying, everything()) 

all_data_gentrify
```

```{r laisha}
all_data_gentrify |> 
  filter(year != 2010) |> 
  pivot_longer(cols = contains('enrollment') & !starts_with('grade') & !starts_with('kinder') & !starts_with('total'), 
               names_to = 'enrollment_group', 
               values_to = 'enrollment_number') |> 
  mutate(pct_total_enrollment = 100*enrollment_number/total_enrollment) |> 
  select(enrollment_group, enrollment_number, total_enrollment, pct_total_enrollment, everything()) |>
  ggplot() + 
  geom_line(aes(x = year, 
             y = enrollment_number, 
             color = enrollment_group)) + 
  scale_color_viridis(discrete = T) + 
  facet_wrap(~enrollment_group) +
  geom_line(aes(x = year, 
                y = total_enrollment), 
            size = 2, 
            color = 'red')
```

```{r verify-gentrification-NEW}
all_data_gentrify |> 
  filter(year != 2010) |> 
  ggplot(aes(x = gentrifying, 
             y = total_enrollment)) + 
  geom_boxplot()


#PIVOTS SECTIONS TO GROUP VARIABLES
all_data_gentrify |> 
  filter(year != 2010) |> 
  pivot_longer(cols = contains('enrollment') & !starts_with('grade') & !starts_with('kinder') & !starts_with('total'), 
               names_to = 'enrollment_group', 
               values_to = 'enrollment_number') |> 
  mutate(pct_total_enrollment = 100*enrollment_number/total_enrollment) |> 
  select(enrollment_group, enrollment_number, total_enrollment, pct_total_enrollment, everything()) |> 
  ggplot(aes(x = enrollment_group, 
             y = pct_total_enrollment)) + 
  geom_boxplot(aes(fill = gentrifying)) +
  coord_flip()

all_data_gentrify |> 
  filter(year != 2010) |>  
  pivot_longer(cols = contains('enrollment') & !starts_with('total') & (starts_with('grade') | starts_with('kinder')), 
               names_to = 'enrollment_group', 
               values_to = 'enrollment_number') |> 
  mutate(pct_total_enrollment = 100*enrollment_number/total_enrollment) |> 
  select(enrollment_group, enrollment_number, total_enrollment, pct_total_enrollment, everything()) |> 
  ggplot(aes(x = enrollment_group, 
             y = pct_total_enrollment)) + 
  geom_boxplot(aes(fill = gentrifying)) +
  coord_flip()

all_data_gentrify |> 
  filter(year != 2010) |> 
  pivot_longer(cols = contains("graduation"), 
               names_to = 'grad_rate_timeframe', 
               values_to = 'grad_rate') |> 
  ggplot(aes(x = grad_rate_timeframe, 
             y = grad_rate)) + 
  geom_boxplot(aes(fill = gentrifying))

all_data_gentrify |> 
  filter(year != 2010) |> 
  pivot_longer(cols = contains("dropout"), 
               names_to = 'dropout_rate_timeframe', 
               values_to = 'dropout_rate') |> 
  ggplot(aes(x = dropout_rate_timeframe, 
             y = dropout_rate)) + 
  geom_boxplot(aes(fill = gentrifying))

all_data_gentrify |> 
  filter(year != 2010) |> 
  pivot_longer(cols = contains('ready'), 
               names_to = 'subject_readiness', 
               values_to = 'pct_ready') |> 
  ggplot(aes(x = subject_readiness, 
             y = pct_ready)) + 
  geom_boxplot(aes(fill = gentrifying)) +
  coord_flip()
```

```{r}
all_data_gentrify |> 
  filter(year != 2010) |>
  pivot_longer(cols = contains('ready'), 
             names_to = 'subject_readiness', 
             values_to = 'pct_ready') |>
  ggplot() +
  geom_tile(aes(x = year,
                y = pct_ready,
                fill = gentrifying), width = 1, height = 1) +
  geom_line(aes(x = year, 
           y = pct_ready)) +
  facet_wrap(~subject_readiness) +
  scale_color_viridis(discrete = T)
```

```{r verify-gentrification-OLD}
#WANT TO HAVE A LINE GRAPH OVER YEAR WITH DIFFERENT LINES FOR ENROLLMENT WITH BACKGROUND COLOR SHADING FOR WHEN A YEAR COUNTS AS BEING GENTRIFIED
all_data_gentrify |> 
  filter(year != 2010) |> 
  pivot_longer(cols = contains('enrollment') & !starts_with('total') & (starts_with('grade') | starts_with('kinder')), 
               names_to = 'enrollment_group', 
               values_to = 'enrollment_number') |> 
  mutate(pct_total_enrollment = 100*enrollment_number/total_enrollment) |> 
  select(enrollment_group, enrollment_number, total_enrollment, pct_total_enrollment, everything()) |> 
  ggplot() +
  geom_tile(aes(x = year, 
             y = pct_total_enrollment, 
             fill = gentrifying), width = 1, height = 1) + 
  scale_fill_brewer(palette = 'Paired') + 
  geom_line(aes(x = year, 
             y = pct_total_enrollment)) + 
  facet_wrap(~enrollment_group)
  scale_color_viridis(discrete = T)
```

# Visualizations

## Verifying Gentrification Part a - Increase in Housing Construction

```{r viz-housing-delta}
all_data_gentrify |> 
  ggplot() + 
  geom_col(data = filter(all_data_gentrify, housing_delta > 0), 
           aes(x = year, 
               y = housing_delta), 
           fill = 'darkgreen') +
  geom_col(data = filter(all_data_gentrify, housing_delta <= 0), 
           aes(x = year, 
             y = housing_delta), 
           fill = 'gray') +
  labs(title = 'Changes in Housing Built in Detroit, Year over Year',
       subtitle = 'Determining Gentrification, Part 1 - Does Housing Construction Increase?', 
       x = 'Year', 
       y = 'Housing Built Since Previous Year') + 
  theme_minimal() 
```

## Verifying Gentrification Part b - Increase in College Educated Households

```{r viz-coll-educated}
all_data_gentrify |> 
  ggplot() + 
  geom_col(data = filter(all_data_gentrify, edu_level_delta > 0), 
           aes(x = year,
               y = edu_level_delta),
            fill = 'darkgreen') +
  geom_col(data = filter(all_data_gentrify, edu_level_delta <= 0), 
           aes(x = year,
               y = edu_level_delta),
            fill = 'gray') +
  labs(title = 'Changes in Educational Attainment, Year over Year', 
       subtitle = 'Determining Gentrification, Part 2 - Is there an Increase in College-Educated Households?',
       x = 'Year', 
       y = "% Change in Education Attainment") + 
  theme_minimal() + 
  annotate("text", 
           label = "Note: 'Educational Attainment' defined\nas having a Bachelor's Degree or Higher",
           x = 2011,
           y = -0.2,
           hjust = 0,
           size = 3,
           fontface = 'italic')

```

```{r viz-ed-attain-demographic}
all_data_gentrify |> 
  pivot_longer(cols = contains('percent_educational_attainment') & !contains("25_years_and_over") & !contains('or_higher'), 
               names_to = 'education_level', 
               values_to = 'percent_at_level') |>
  mutate(education_level = str_replace_all(education_level, "percent_educational_attainment_", ""), 
         education_level = str_replace_all(education_level, "9th_to_12th_grade_no_diploma", "Some High School"), 
         education_level = str_replace_all(education_level, "associates_degree", "Associate's Degree"), 
         education_level = str_replace_all(education_level, "bachelors_degree", "Bachelor's Degree"), 
         education_level = str_replace_all(education_level, "graduate_or_professional_degree", "Graduate Degree"), 
         education_level = str_replace_all(education_level, "high_school_graduate_includes_equivalency", "High School Graduate"),
         education_level = str_replace_all(education_level, "less_than_9th_grade", "Less than 9th Grade"), 
         education_level = str_replace_all(education_level, "some_college_no_degree", "Some College")) |>
  filter(education_level == "Associate's Degree" | education_level == "Bachelor's Degree" | education_level == "Graduate Degree" | education_level == "Some High School") |> 
  ggplot() + 
  geom_col(aes(x = year,
               y = percent_at_level, 
               fill = factor(education_level,
                    ordered = TRUE,
                    levels = c("Associate's Degree", 
                               "Some High School",
                               "Bachelor's Degree",
                               "Graduate Degree")))) +
  facet_wrap(~education_level, ncol = 4) +
  theme_minimal() + 
  theme(legend.position = 'none') +
  scale_fill_manual(values = c('darkgreen', 'darkgreen', 'darkgreen', 'darkgreen')) + 
  labs(title = "Population Education-Level Demographic Change, 2010-2018",
       subtitle = 'Less High School Droups and More College Graduates',
       x = "Year", 
       y = "% of Population at Education Level")
```

```{r}
#NOT SURE WE NEED THIS
all_data_gentrify |> 
  pivot_longer(cols = contains('percent_educational_attainment') & !contains("25_years_and_over") & contains('or_higher'), 
               names_to = 'education_level', 
               values_to = 'percent_at_level') |> 
  mutate(education_level = str_replace_all(education_level, "percent_educational_attainment_percent_", ""),
         education_level = str_replace_all(education_level, "high_school_graduate_or_higher", "Diploma/GED or Higher"), 
         education_level = str_replace_all(education_level, "bachelors_degree_or_higher", "Bachelor's Degree or Higher")) |>
  ggplot() + 
  geom_col(aes(x = year,
                y = percent_at_level), 
           fill = 'darkgreen') + 
  facet_grid(~education_level)
```

## Verifying Gentrification Part c - Increase in Housing Costs

```{r viz-housing-cost}
p1 <- all_data_gentrify |> 
  ggplot() + 
  geom_col(data = filter(all_data_gentrify, rent_delta > 0), 
           aes(x = year,
               y = rent_delta),
            fill = 'darkgreen') +
  geom_col(data = filter(all_data_gentrify, rent_delta <= 0), 
           aes(x = year,
               y = rent_delta),
            fill = 'gray') +
  labs(x = 'Year', 
       y = "Change in Median Rent") + 
  theme_minimal()

p2 <- all_data_gentrify |> 
  ggplot() + 
  geom_col(data = filter(all_data_gentrify, value_delta > 0), 
           aes(x = year,
               y = value_delta),
            fill = 'darkgreen') +
  geom_col(data = filter(all_data_gentrify, value_delta <= 0), 
           aes(x = year,
               y = value_delta),
            fill = 'gray') +
  labs(x = 'Year', 
       y = "Change in Median Housing Value") + 
  theme_minimal()

p1 + p2 + plot_annotation(title = 'Changes in Housing Costs, Year over Year', 
                          subtitle = 'Determining Gentrification, Part 3 - Is there an Increase in the Cost of Housing?')
```

```{r}

all_data_gentrify |> 
  mutate(pct_econ_disadv_enrollment = 100*economic_disadvantaged_enrollment/total_enrollment) |> 
  ggplot(aes(x = gross_rent_median_dollars, 
             y = pct_econ_disadv_enrollment)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) +
  geom_text_repel(aes(label = year))

all_data_gentrify |> 
  mutate(pct_econ_disadv_enrollment = 100*economic_disadvantaged_enrollment/total_enrollment) |> 
  #mutate(year = as.character(year)) |>
  ggplot(aes(x = year, 
             y = gross_rent_median_dollars, 
             size = pct_econ_disadv_enrollment)) + 
  geom_point() 
```

```{r}
all_data_gentrify |> 
  mutate(pct_econ_disadv_enrollment = 100*economic_disadvantaged_enrollment/total_enrollment) |> 
  ggplot(aes(x = value_median_dollars, 
             y = pct_econ_disadv_enrollment)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) +
  geom_text_repel(aes(label = year))

all_data_gentrify |> 
  mutate(pct_econ_disadv_enrollment = 100*economic_disadvantaged_enrollment/total_enrollment) |> 
  ggplot(aes(x = value_median_dollars, 
             y = pct_econ_disadv_enrollment)) + 
  geom_point(aes(size = total_enrollment)) + 
  geom_smooth(method = 'lm', se = FALSE) +
  geom_text_repel(aes(label = year))

all_data_gentrify |> 
  mutate(pct_econ_disadv_enrollment = 100*economic_disadvantaged_enrollment/total_enrollment) |> 
  mutate(year = as.character(year)) |>
  ggplot(aes(x = year, 
             y = value_median_dollars, 
             size = pct_econ_disadv_enrollment)) + 
  geom_point() 
```

```{r relationship-testing}
all_data_gentrify |> 
  
  ggplot(aes(x = gross_rent_median_dollars, 
             y = value_median_dollars)) + 
  geom_point() + 
  geom_smooth(method = 'lm', 
              se = FALSE)

all_data_gentrify |> 
  mutate(year = as.character(year)) |> 
  ggplot() + 
  geom_point(aes(x = year, 
                 y = gross_rent_median_dollars, 
                 size = total_enrollment)) 

all_data_gentrify |> 
  mutate(year = as.character(year)) |> 
  ggplot(aes(x = total_enrollment, 
                 y = gross_rent_median_dollars)) + 
  geom_point() + 
  geom_smooth(method = 'lm') +
  geom_text_repel(aes(label = year))
```
